\input{preamble.tex}
\title{Periodic Functions with Sum as Identity Function}
\created{30 Jan 2019}
\input{begin.tex}
\section*{Problem}
Find two periodic functions \( f \) and \( g \) from \( \mathbb{R} \)
to \( \mathbb{R} \) such that their sum \( f + g \) is the identity
function. The axiom of choice is allowed.

A function \( f \) is periodic if there exists \( p > 0 \) such that
\( f(x + p) = f(x) \) for all \( x \) in the domain.

\section*{Solution}
The axiom of choice is equivalent to the statement that every vector
space has a basis. Since the set of real numbers \( \mathbb{R} \) is a
vector space over the set of rational numbers \( \mathbb{Q} \), there
must be a basis \( \mathcal{H} \subseteq \mathbb{R} \) such that every
real number \( x \) can be written uniquely as a finite linear
combination of elements of \( \mathcal{H} \) with rational coefficients,
i.e.,

\[
    x = \sum_{a \in \mathcal{H}} x_a a
\]

where each \( x_a \in \mathbb{Q} \) and \( \{ a \in \mathcal{H} \mid x_a
\ne 0 \} \) is finite. The set \( \mathcal{H} \) is also known as the
Hamel basis.

We know that \( b_a = 0 \) for distinct \( a, b \in \mathcal{H} \)
because \( a \) and \( b \) are basis vectors. In the above expansion of
\( x \), each \( x_a \) is a rational number that appears as the
coefficient of the basis vector \( a \). Therefore \( (x + y)_{a} = x_a
+ y_a \) for all \( x, y \in R \). Thus \( (x + b)_{a} = x_a + b_a = x_a
+ 0 = x_a \). This shows that a function \( f(x) = x_a \) is a periodic
function with period \( b \) for any \( b \in \mathcal{H} \setminus
\{a\} \).

Let us define two functions:
\[
    f(x) = \sum_{a \in \mathcal{H} \setminus \{ b \}} x_a a,
    \hspace{1em}
    g(x) = x_b b.
\]

where \( b \in \mathcal{H} \) and \( x \in \mathbb{R} \). Let us
choose \( c \in \mathcal{H} \) such that \( c \ne b \). Then \( f(x) \)
is a periodic function with period \( b \) and \( g(x) \) is a periodic
function with period \( c \). Further,

\[
f(x) + g(x)
= \left( \sum_{a \in \mathcal{H} \setminus \{ b \}} x_a a \right) + x_b b
= \sum_{a \in \mathcal{H}} x_a a
= x.
\]

Thus \( f(x) \) and \( g(x) \) are two periodic functions such that
their sum is the identity function.

\nocite{pfml-periodic-functions}
\nocite{radcliffe-periodic-functions}
\nocite{youcis-dim-r-q}
\input{end.tex}
